### 2020-05-21 paper.tex
Edit
Comparison

We could verify previously gathered interview data with the results from the new reports: The subjects tend to manually add log statements to their code in order to understand its behavior, especially when things get more complex. Three subjects even claimed using specialized logging functionalities provided by rxjs-spy\cite{rxjsspy}. As the subjects stated, the knowledge about additional events made visible helped in situations where they needed to find problems within a graph composing multiple asynchronous data sources on multiple levels. Without it, they would have been blind to what actually happened within the system.

@@

\begin{listing}[H]
	\begin{minted}{TypeScript}
		interval(1000).pipe(      // Increasing numbers with 1s delay
		  map(i => i * 2),        // Breakpoint on i * 2
		  map(i => i - 1)
		).subscribe(console.log); // -1, 1, 3, 5, 7, ...
	\end{minted}
	\caption{Example RxJS Stream With Breakpoint on Line 2 Within the Arrow Function}
	\label{lst:rxjs-step-controls}
\end{listing}

@@

Our work is not the first approach on understanding and improving on the struggles software engineers have to overcome when working on RP code, either with or without RxJS. We are going to present four of this tools in this section. The list is not meant to be exhaustive. It contains tools we either got to know during our research, because they were mentioned in one of the interviews or we worked with them personally in the past.

@@

@@

@@

\begin{listing}[H]
	\begin{minted}{TypeScript}
		interval(1000).pipe(
		  map(i => i * 2),
		  tag('multiply'),        // Create tag "multiply"
		  map(i => i - 1)
		).subscribe(console.log);
	\end{minted}
	\caption{Usage of \emph{rxjs-spy} \mintinline{TypeScript}{tag} Operator}
	\label{lst:rxjs-spy-tag}
\end{listing}

@@

@@

>>>>

We could verify previously gathered interview data with the results from the new reports: The subjects tend to manually add log statements to their code in order to understand its behavior, especially when things get more complex. Three subjects even claimed using specialized logging functionalities provided by rxjs-spy\cite{rxjsspy}. As the subjects stated, the knowledge about additional events made visible helped in situations where they needed to find problems within a graph composing multiple asynchronous data sources on multiple levels. Without it, they would have been blind to what actually happend within the system.

@@

\begin{listing}[H]
	\begin{minted}{TypeScript}
		interval(1000).pipe(      // Increasing numbers with 1s delay
		  map(i => i * 2),        // Breakpoint on i * 2
		  map(i => i - 1),
		  take(2)
		).subscribe(console.log); // -1, 1
	\end{minted}
	\caption{Example RxJS Stream With Breakpoint on Line 2 Within the Arrow Function}
	\label{lst:rxjs-step-controls}
\end{listing}

@@

Our work is not the first approach on understanding and improving on the struggles software engineers have to overcome when working on RP code, either with or without RxJS. We are going to present four of this tools in this section. This list is not meant to be exhaustive. It contains tools we either got to know during our research because they were mentioned in one of the interviews or we worked with them personally in the past.

@@

Salvaneschi et al. claim\cite{10.1145/2577080.2577083} the architecture behind \emph{Reactive Inspector} to be flexible enough to be used with different RP runtime environment as well.

@@

The architectural design of \emph{rxfiddle} is similar to the one \emph{Reactive Inspector}: Banken et al. provide a system where different RP runtime environments should be able to be attached to the Visualizer easily.

@@

\begin{listing}[H]
	\begin{minted}{TypeScript}
		interval(1000).pipe(
		  map(i => i * 2),
		  tag('multiply'),
		  map(i => i - 1),
		  take(2)
		).subscribe();
	\end{minted}
	\caption{Usage of \emph{rxjs-spy} \mintinline{TypeScript}{tag} Operator on Line 3}
	\label{lst:rxjs-spy-tag}
\end{listing}

@@

\begin{listing}[H]
	\begin{minted}{Typescript}
		Tag = multiply; notification = subscribe
		Tag = multiply; notification = next; value = 0
		Tag = multiply; notification = next; value = 2
		Tag = multiply; notification = unsubscribe
	\end{minted}
	\caption{Trace log generated by \emph{rxjs-spy} \mintinline{TypeScript}{tag} from Listing~\ref{lst:rxjs-spy-tag}}
	\label{lst:rxjs-spy-log}
\end{listing}

@@

\subsection{Comparison}

Specialized utilities are aware of RP runtime environments and help software engineers to interpret its state better than before using sole traditional debugger tools. Most of them provide a way to represent the data flow graph either graphically (Reactive Inspector, rxfiddle) or in textual form (rxjs-spy). They visualize the data flowing through the graph over time in marble diagrams (rxfiddle, Rx Visualizer) or directly in the visual graph representation (Reactive Inspector). Some like Reactive Inspector for REScala provide an ``all in one'' solution with breakpoints, omniscient debugging capabilities and full IDE integration whereas the RxJS-based tools have a more specific feature set: They usually work outside of an IDE and/or need manual code instrumentation like rxjs-spy.

\textbf{TODO Compare im Table?}
